Identification of Direct Cherenkov Pixels using Boosted Decision Trees

### 2016-07-06 report/final.tex
Post-Thesis Backup 06/07/16

\documentclass[11pt]{article}
\usepackage{graphicx}
\graphicspath{ {graphs/} }
\usepackage{geometry}
\geometry{a4paper, portrait, margin=1.0in}
\usepackage{gensymb}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{pgfplotstable, booktabs}
\usepackage{caption}
\usepackage{subcaption}

@@

\begin{document}
\begin{titlepage}
    \begin{center}
        \vspace*{1cm}

@@

    \end{center}
\end{titlepage}
\section*{Abstract}
Between impact with the upper atmosphere and decay into a charged particle shower, heavy cosmic ray elements such as Iron emit Cherenkov Light at an angle determined by the Refractive Index of the air and the energy per nucleon. This direct Cherenkov Light forms a characteristic circular light distribution on the Earth's surface with an intensity proportional to the square of the cosmic ray charge. A new method has been developed to reconstruct this charge number, by fitting the received Cherenkov Photons to the characteristic Lateral Photon Distribution. The expected performance for various existing and planned installations will be discussed.

\section*{Zusammenfassung}
Between impact with the upper atmosphere and decay into a charged particle shower, heavy cosmic ray elements such as Iron emit Cherenkov Light at an angle determined by the Refractive Index of the air and the energy per nucleon. This direct Cherenkov Light forms a characteristic circular light distribution on the Earth's surface with an intensity proportional to the square of the cosmic ray charge. A new method has been developed to reconstruct this charge number, by fitting the received Cherenkov Photons to the characteristic Lateral Photon Distribution. The expected performance for various existing and planned installations will be discussed.
\newpage
\tableofcontents
\newpage

\section{Preface}
The basis of this thesis work can be divided into two halves. The first half is the simulation of cosmic ray HESS camera images, and analysis work based on these images. The second half is a Monte Carlo study simulating air shower intensities, based entirely on function fitted from the HESS camera images.

It is important to note that the simulation of cosmic ray air showers for the first part of my thesis was performed using the CORSIKA software \cite{Heck98}, and the resultant HESS camera images were produced using the Sim\textunderscore Tel array software \cite{Bernlohr08}. I claim no credit for the complex simulations that these packages have enabled me to perform.

@@

Aside from this, all other work in this thesis is my own. I wrote all necessary code for the Monte Carlo simulation, and it was conducted entirely by me. Similarly

Thank Dieter, and Attilla, and...
\newpage

\section{Introduction}
Cosmic Rays are abundant in our solar system, entering the Earth's upper atmosphere from every direction. Cosmic Rays are in fact not rays but rather charged particles, primarily protons, which have been accelerated by various galactic or stellar processes. Their flux is governed by a well-defined energy power law, which is almost unchanged across more than 10 orders of magnitude \cite{Blasi:2013rva}.  The exact Cosmic Ray composition is energy dependent, and provides information regarding their sources. Heavy Cosmic Rays, the most common of which is iron, vary substantially in relative abundance for different energy scales.

@@

To fully reconstruct an event, we will need to find the charge, energy and the core position coordinates. In addition, we will need to determine the height of the first atmospheric interaction. We would normally require at least five data points in order to reconstruct these variables. However, it is possible to use the non-observation of light in one telescope to constrain the core position. Thus, we require that an event was be seen by at least four telescopes for a five-telescope array such as HESS. An illustration of the HESS array is shown in Figure \ref{fig:hessdiagram}.

@@

The refractive index as a function of height is shown in as seen in Figure \ref{fig:lpd}, using standard values listed in the appendix. Because the refractive index of air increases very quickly as the height decreases, the Cherenkov Angle $\theta_{C}$ will also increase rapidly with decreasing height. Thus, although the displacement $(h - altitude_{array})$ decreases as the height, the radius on the ground increases as height decreases. Consequently the upper atmosphere emission contributes to the inner LDF, while the lower atmosphere emission contributes to the outer LDF. As the cosmic ray travels down through the atmosphere, Cherenkov Emission continuously illuminates new areas on the ground that were previously unlit, without re-illuminating inner regions.

Emission continues until the first interaction with the atmosphere, occurring at a randomly distributed height we call $h_{int}$. The maximum radius of the DC light will then be $r_{max} = Radius(h_{int})$. This is extremely useful, because for all distances up to $r_{max}$, the LPD will be completely independent of the emission height. One full parameterisation of the LPD will thus be valid for all heights, provided the maximum radius is known.

@@

A full simulation of the LPD was undertaken, with the assumption that $\Delta_{E} \approx 0.7$ independent of Energy. Although this is certainly an oversimplification of the true picture, it will allow us to understand the variation of the LPD with height. A more rigorous LPD parameterisation will be found later to account for the dependence of $\Delta_{E}$ on the Energy.

If we divide the number of emitted photons by the area of the annulus between the ring radii at $h$ and $h - \Delta_{h}$, we retrieve the LDF shown in \ref{fig:lpd}, which varies with $ \rho_{DC}  = f(r) \times Z^{2}$ . Thus the amplitude of the LDF is proportional to the square of the Cosmic Ray charge, enabling us to reconstruct this number from the DC emission. This proportionality is the basis for charge reconstruction in the LDF method.

\begin{figure}
\begin{center}
\includegraphics[height=0.9\textheight]{logenergyradius}
\caption{The Threshold Energy for Cherenkov Emission is marked in blue. With the assumption of $\beta=1$, the maximum emission radius is marked in black. The red and green and magenta line show the emission radius at 3.57 and 0.75 and 0.23 TeV per Nucleon respectively. The Green line is sufficiently close to the background to be saturated at 24km, while the magenta line is not.}
\label{fig:generalenergy}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=0.9\textheight]{simulatedlpd}
\caption{Above, the refractive index is shown with both linear and exponential interpolation between the known values. The LPD obtained from simulation of an Iron Nucleus up to a first interaction height of 25km for a range of Core Energies. An altitude of 1.8km for the experimental array is assumed. Atmospheric absorption, although neglected, is broadly constant across the emission range leading to uniform amplitude scaling.}
\label{fig:lpd}
\end{center}
\end{figure}

The refractive index at a series of heights, based on data from the HESS site, is shown \ref{fig:lpd}. Exponentials are fitted between points to provide interpolation, and a comparison can be made to a linear interpolation that is also plotted. Despite the exponential interpolation, discontinuities in the second derivative of the refractive index prevent the LPD from being smooth in the simulation. This is unimportant, because in reality the variation due to atmospheric conditions and random noise will smear out any discontinuities in the LDF.

@@

\subsection{Classic HESS1 DC Pixel Identification}
We can find the DC pixel in am image by defining the variable $Q_{DC} = \frac{Intensity_{N.N.max}}{Intensity}$ as the ratio of the largest neighbouring pixel intensity to the intensity of a given pixel. In previous experiments, the DC pixel candidate was identified by applying a number of cuts to pixels in an image \cite{hess07}, and from the subset of pixels passing the specified cuts, selecting the pixel with the smallest $Q_{DC}$ as the \textquoteleft DC candidate'. Due to the low pass rate for cuts, we obtain a small low-contamination dataset, while the majority of telescope images are left without a DC pixel candidate.

As a basis for comparison, these original HESS cuts listed in Table \ref{tab:qdccuts} were replicated for the set of test data. For every HESS1 image, the total image amplitude $I_{tot}$ was used alongside the zenith angle $\theta$ to determine a dynamic cut, $Q_{DC} < 0.14 \times \log(\frac{I_{tot}}{161 \times \cos \theta})$. Because many images had no pixel that passed all cuts, the $Q_{DC}$ method was frequently unable to identify a DC pixel. In the original analysis, an additional cut $r_{core} \textgreater 40m$ was applied. However, the uncertainty in determining the core position through Hillas Analysis is typically of the order of $\pm 30m$. Consequently, this particular cut was omitted.

@@

In general, larger training sets prevent overtraining by making it more difficult to uniquely distinguish the individual events. The maximum depth of a tree is the number of branches a decision chain can pass through before it must end on a leaf. The minimum leaf sample is the minimum number that every leaf has sample must exceed. Either of these quantities can be used to restrict overtraining of a Decision Tree.

@@

The results are summarised in Table \ref{tab:qdcbdtcomparison1}. The BDT method represents a clear improvement in DC pixel identification over the previous $Q_{DC}$ method, and corresponds to a fifteen-fold increase in identified pixels.

@@

\subsection{HESS-2 BDT}
The original HESS study predated the construction of the larger CT5 telescope, and thus focused exclusively in the four HESS-1 telescopes. Although the cuts in Table \ref{tab:qdccuts} were not optimised for the differing pixel size and angular viewing region of CT5, they were replicated to provide a basis for comparison. Based on CT5 images from the training sample, an increased 5.6\% of DC pixels were correctly identified and passed the required cut. A further 11.3 \% of all pixels passed the cuts, despite being misidentified. This represented a very heavy decrease in sample purity to just 33.0\%. We can assume that better optimised CT5 cuts could remove many of these misidentified events, though it is unlikely that any significant improvement in the number of correctly identified DC pixels would be possible. Thus, the number of DC pixels successfully identified using the $Q_{DC}$ method is still a valid benchmark for comparative BDT performance. A true accuracy rate closer to the HESS 1 rate of 90\% could be expected, at the expense of a depressed acceptance rate. Due to the distinctiveness of the CT5 telescope hardware, direct application of the HESS1 classifier to the CT5 telescope yields poorer BDT performance. Discrepancies in To improve CT5 pixel identification, a separate HESS2 BDT was instead trained with the same variables as above. The relative feature importance is listed in Table \ref{tab:hess2classifier}.

@@

Despite the minor performance difference, the BDT method is clearly superior in both cases. On top the gains in BDT accuracy when considering only high-multiplicity events high multiplicity events, we also find the relative performance gap over the $Q_{DC}$ method is vastly increased. As a result of the two-hundred-fold increase in acceptance rate for high-multiplicity-BDT events against high-multiplicity-$Q_{DC}$ events, the expected data sample size will increase dramatically. This should enable spectroscopic analysis of Cosmic Ray elements to be conducted after a much shorter experimental run time that for the $Q_{DC}$ method. Use of the BDT identification method will be assumed throughout the rest of this paper, and a high-multiplicity event will be intended to mean a high-multiplicity-BDT event.

\begin{table}[h!]
  \centering
  \caption{Comparison of high-multiplicity performance Performance}
  \label{tab:highmultiplicitycomparison}
  \begin{tabular}{c|cc|cc}
    \toprule
    & \multicolumn{2}{c|}{HESS 1} & \multicolumn{2}{c}{HESS 2} \\
    & $Q_{DC}$ & BDT & $Q_{DC}$ & BDT\\
    \midrule
    High Multiplicity Accepted Pixels Correctly Identified (\%)& 0.01 & 2.20 & 0.01 & 1.61\\
    High Multiplicity Accepted Pixels Incorrectly Identified (\%)  & 0.00 & 0.43 & 0.01 & 0.22\\
    Sample Purity (\%)& 100.0 & 83.8 & 100.0 & 88.0\\
    \bottomrule
  \end{tabular}
\end{table}

@@

We require an additional cut to remove the majority of these proton events. If we can eliminate one or two accepted proton images per event, then the remaining proton pixels will all be rejected by the multiplicity cuts. Hence, loose cuts can sharply reduce the acceptance rate of protons. In this case, as shown in Figure \ref{fig:aspectratio}, there is a fair degree of separation between proton and Iron Aspect Ratio for pixels which pass all cuts. We aim to remove proton pixels but not iron pixels. Applying a cut requiring Aspect Ratio $> 0.40$ eliminated all accepted proton events for HESS1 and HESS2. With 17792 total HESS1 images, just 6195 images were triggered by DC light. Of all 17792 images, there were none passing all of the required cuts.

@@

Thus, if we have thirtyfold increase in event rate, but only one-fifteenth the acceptance rate, we would ultimately expect that we would have twice as many Iron Events if the full energy range of 5-135 TeV was included. However, the low-energy proton acceptance is much higher. Thus, were we to include all events in this energy range, we would again retrieve a sample dominated by protons. The results are summarised in Table \ref{tab:fullironacceptance}.

@@

To remove the proton contamination from protons and non-saturated Iron events, we thus apply one additional Energy cut on the data sample. The majority of accepted protons have an Energy of less than 10 TeV, and the largest recorded Energy was 21.6 TeV. We can thus require that the reconstructed Energy is greater than 30TeV. This will, assuming a rough energy resolution of around 15\% in reconstruction, easily remove all of the proton background. We will also remove roughly half of the Iron events, but as shown in Figure \ref{fig:lpd}, this low-energy half of events will mostly contain non-saturated DC emission that would be harder to reconstruct.

@@

\section{Parameterisation of the LPD}
Having identified DC pixels in a reliable way without accompanying background events, we can now consider the resultant LPD. We must parameterise this function before we are able to reconstruct events. We can also account for additional information provided in the EAS light.

\subsection{Full-Shower LPD}
Having restricted ourselves to high-multiplicity events, we reduce our data sample to mostly images in which air shower lies relatively close to the center of the telescope field of view. We can conclude that in many of our images, the air shower will lie mostly within the telescope image. In the same way as for DC light, we can firstly look for the existence of a Characteristic LPD describing the EAS Intensity, where we define $I_{tot}$=Image Amplitude. The results for high-multiplicity events at various energies is shown in Figure \ref{fig:fullshowerlpd}.

@@

For each energy, a clear exponential decay in Image Amplitude is observed. The data is fitted with a line of the form $y_{i} (x) =A_{i} \exp (K_{i} x)$. The fractional deviation from the fit is found for each point by calculating $\Delta_{EAS} = \frac{fit_{EAS} - I_{tot}}{fit_{EAS}}$. We can then take the 68\% interval values for fractional deviation to give a fractional uncertainty $\sigma_{EAS}$. Across all Energies, the mean fractional deviation from the fitted exponential was roughly 0.15. For the purposes of reconstruction, an exponential decay with a varying exponent and amplitude will not be particularly useful, because there are two degrees of freedom.

@@

Having retrieved a law for K, each dataset is then fitted a second time. The exponent is determined by the Energy, leaving only one degree of freedom for the fit. The fitted Amplitudes $A_{i}$ are then recorded, and are also plotted in Figure \ref{fig:powerlawlpd}. A clear relationship is again observed, in this an exponential law. A straight line is fitted to the logarithm of the data values, and we retrieve a second relationship describing the amplitude as a function of the Energy.

@@

Through combined use of the two equations, we can parameterise the entire Full-Shower LPD using only the Energy. These fitted LPDs are also plotted in Figure \ref{fig:fullshowerlpd}. The fractional deviation bands are also shown in green, with a mean fractional deviation of roughly 0.15 across the energies. The Full Shower LPDs continue to be large enough to trigger telescopes for several hundred meters. It is thus likely that, for high-multiplicity events, we will obtain an additional four or five data points from the Full Shower LPD.

@@

Although the Full Shower LPDs do not depend upon the charge of the primary particle, they will still enable the core position and energy to be effectively constrained. We can use the Full Shower LPD to obtain a rough estimate for the core position, and thus a rough estimate of the distance from each telescope to the core. Similarly the core energy can be estimated. Although this reconstruction will be inferior to one accounting for both LPDs, we can make use of the preliminary core position estimate to aid our calculation of $True_{DC}$. This in turn should reduce the uncertainty when determining the charge of the primary particle.

@@

\subsection{Determining $True_{DC}$}
Before we can extract the DC LPD from telescope imagesTo determine the value of $True_{DC}$ is essential for determining the error in our LPD measurements. However, there are two natural ways of quantifying $True_{DC}$. In the basic case, we can simply consider the EAS-free intensity $Intensity_{max}$ in the DC pixel. An alternative is to take the entire EAS-free image amplitude $I_{tot}$ as our $True_{DC}$ value. The use of Image Amplitude accounts for the fact that the DC light is almost always split between two or more pixels. Although in most cases, one pixel has the majority of the light, in extreme cases the light can be evenly split between two. As a result of the smaller angular region covered by each CT5 pixel, this happens much more frequently for HESS2 than for HESS1. A countervailing argument is the susceptibility of the Image Amplitude measurement to edge effects. A DC pixel lying near the image edge will be unaffected by this fact, but the total image amplitude may decline because dimmer pixels move over the image edge. We hope that, as with the full-shower LPD measurements, restricting ourselves to high-multiplicity events will enable us to remove most images that would otherwise suffer from edge effects.

>>>>

\documentclass[11pt]{article}
\usepackage{graphicx}
\graphicspath{ {graphs/} }
\usepackage{geometry}
\geometry{a4paper, portrait, margin=1.0in}
\usepackage{gensymb}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{pgfplotstable, booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage{afterpage}
\newcommand\blankpage{%
    \null
    \thispagestyle{empty}%
    \addtocounter{page}{-1}%
    \newpage}
\onehalfspacing

@@

\begin{document}
\afterpage{\blankpage}
\begin{titlepage}
    \begin{center}
        \vspace*{1cm}

@@

    \end{center}
\end{titlepage}
\section*{Abstract}
Between entry into the upper atmosphere and interaction with the lower atmosphere to produce a charged particle shower, heavy Cosmic Ray elements such as Iron will emit Cherenkov Light. This Direct Cherenkov (DC) Light forms a characteristic circular light distribution on the Earth's surface with an intensity proportional to the square of the cosmic ray charge. While the air showers will cover much of a telescope image, this DC light is usually concentrated in a single DC pixel. A new method of identifying these DC pixels was developed using a Boosted Decision Tree (BDT), leading to an eightfold increase in acceptance rate and a decrease in misidentification. A second new method was developed to reconstruct this charge number, by fitting the received Cherenkov Photons to the characteristic Lateral Photon Distribution (LPD). The reconstruction requires at least four DC pixels to be found simultaneously, with the new BDT identification method leading to a two-hundred-fold increase in such events being found. A likelihood minimisation enabled successful reconstruction of the Charge Z of such events. A charge resolution of $\sigma_{Z} \approx 1$ was observed, compared to a resolution of  $\sigma_{Z} \approx 5$ that existing techniques are capable of.

\section*{Zusammenfassung}
Nach dem Eintritt in den oberen Atmosph\"{a}re und vor die Interaktion mit den unteren Atmosph\"{a}re, die eine Teilchenkaskade produziert, emmitiert schwere Kosmischiche Strahlung wie Eisern das Cherenkov-Licht. Dieses Direkt Cherenkov (DC) Licht produziert eine charakteristische kreisformige lichtkurze auf die Erdefl\"{a}che, die eine Intensit\"{a}t proportional zur Kernladungszahl zum Quadrat hat. Obwohl die Teilchenkaskade die Meisten des Telescopbilds besetzt, dieses Cherenkov-Licht wird normalerweise in einem einzelen Bildpunkt konzentriert. Ein neues Verfahren, das diesen DC-Bildpunkten mit einem Boosted Decision Tree (BDT) identifiziert, ist entwickelt. Es hat nicht nur zu einen achtmaligen Zunehmung der Akzeptiertenquote, sondern auch zu eine Verbessung der Misidentifikationsquote gef\"{u}hrt. Noch ein neues Verfahren, das die Ladungszahl durch einem Fit des Cherenkov-Lichts zum characteristische LPD rekonstruktiert, ist entwickelt. Solche Rekonstruktion braucht gleichzeitig mindestens 4 DC-Bildpunkten, die das neues BDT-Verfahren zeihundertmal so oft finden kann. Dadurch ist eine W\"{a}hrscheinlichkeitsminimierung
m\"{o}glich, die zum Rekonstruktion der Laudungszahl f\"{u}hrt. Letztendlich beobachten wir eine Ladungszahlaufl\"{o}sung von $\sigma_{Z} \approx 1$, die vergleichbar mit den heutigen Aufl\"{o}sung von $\sigma_{Z} \approx 5$ ist.
\newpage
\tableofcontents
\newpage

\section{Preface}
This analysis was conducted in the Astroteilchenphysik group of Universit\"{a}t Hamburg between October 2015 and July 1016. It was supervised by Professor Dieter Horns who, in addition to invaluable guidance and never-ending patience, proposed both the original aim of the project and the concept of LPD reconstruction. It was further supported by Attila Abramowski, whose overlapping field of research provided invaluable opportunities for discussion and comparison.

The basis of this thesis work can be divided into two halves. The first half is the simulation of Cosmic Ray HESS camera images, and analysis work based on these images. The second half is a Monte Carlo study simulating air shower intensities, based entirely on function fitted from the HESS camera images. It is important to note that the simulation of Cosmic Ray air showers for the first part of my thesis was performed using the CORSIKA software \cite{Heck98}, and the resultant HESS camera images were produced using the Sim\textunderscore Telarray software \cite{Bernlohr08}. I claim no credit for the complex simulations that these packages have enabled me to perform.

@@

Furthermore, a part of this analysis was based on comparison with existing methods for reconstructing Cosmic Ray Charge Number that were performed by the HESS collaboration \cite{hess07}. Their groundwork based on the first phase of the HESS array was both a guide and a standard for my efforts at reconstruction.

Aside from this, all other work described here is my own. In particular, I wrote scripts to read the Sim\textunderscore Telarray output, convert raw voltages to calibrated pixels Intensities, and to plot the HESS camera images such as that found in Figure \ref{fig:DCtelimage}. Furthermore, I wrote all necessary code for the Monte Carlo simulation, and it was conducted solely for this analysis. 
\newpage

\section{Introduction}
Cosmic Rays are abundant in our solar system, entering the Earth's upper atmosphere from every direction. Cosmic Rays are in fact not rays but rather charged particles, primarily protons, which have been accelerated by various galactic or stellar processes. Their flux is governed by a well-defined energy power law, which is almost unchanged across more than 10 orders of magnitude \cite{Blasi:2013rva}.  The exact Cosmic Ray composition is energy dependent, and provides information regarding their sources. Heavy Cosmic Rays, the most common of which is Iron, vary substantially in relative abundance for different energy scales.

@@

To fully reconstruct an event, we will need to find the charge, energy and the core position coordinates. In addition, we will need to determine the height of the first atmospheric interaction. We would normally require at least five data points in order to reconstruct these variables. However, it is possible to use the non-observation of light in one telescope to constrain the core position. Thus, we will seek events seen by at least four telescopes for a five-telescope array such as HESS. An illustration of the HESS array is shown in Figure \ref{fig:hessdiagram}.

@@

The refractive index as a function of height is shown in as seen in Figure \ref{fig:lpd}, using standard values listed in the appendix. Because the refractive index of air increases very quickly as the height decreases, the Cherenkov Angle $\theta_{C}$ will also increase rapidly with decreasing height. Thus, although the displacement $(h - altitude_{array})$ decreases with height, the radius on the ground will nonetheless increase as height decreases. Consequently the upper atmosphere emission contributes to the inner LPD, while the lower atmosphere emission contributes to the outer LPD. As the cosmic ray travels down through the atmosphere, Cherenkov Emission continuously illuminates new areas on the ground that were previously unlit, without re-illuminating inner regions.

Emission continues until the first interaction with the atmosphere, occurring at a randomly distributed height we call $h_{int}$. The maximum radius of the DC light will then be $r_{max} = Radius(h_{int})$. This is extremely useful, because for all distances up to $r_{max}$, the LPD will be completely independent of the interaction height. One full parameterisation of the LPD will thus be valid for all heights, provided the maximum radius is known.

@@

A full simulation of the LPD was undertaken, with the assumption that $\Delta_{E} \approx 0.7 eV$ independent of Energy. Although this is certainly an oversimplification of the true picture, it will allow us to understand the variation of the LPD with height. A more rigorous LPD parameterisation will be found later to account for the dependence of $\Delta_{E}$ on the Energy.

If we divide the number of emitted photons by the area of the annulus between the ring radii at $h$ and $h - \Delta_{h}$, we retrieve the LPD shown in Figure \ref{fig:lpd}, which varies with $ \rho_{DC}  = f(r) \times Z^{2}$ . Thus the amplitude of the LPD is proportional to the square of the Cosmic Ray charge, enabling us to reconstruct this number from the DC emission. This proportionality is the basis for charge reconstruction in the LPD method.

\begin{figure}
\begin{center}
\includegraphics[height=0.9\textheight]{logenergyradius}
\caption{The Threshold Energy for Cherenkov Emission is marked in blue. With the assumption of $\beta=1$, the maximum emission radius is marked in black. The red and green and magenta line show the emission radius at 3.57 and 0.75 and 0.23 TeV per Nucleon respectively. The Green 0.7 TeV line is sufficiently close to the background to be saturated at 24km, while the magenta line is not.}
\label{fig:generalenergy}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[height=0.9\textheight]{simulatedlpd}
\caption{Above, the refractive index is shown with both linear and exponential interpolation between the known values. The LPD obtained from simulation of an Iron Nucleus up to a first interaction height of 20km for a range of Core Energies. An altitude of 1.8km for the experimental array is assumed. Atmospheric absorption, although neglected, is broadly constant across the emission range leading to uniform amplitude scaling.}
\label{fig:lpd}
\end{center}
\end{figure}

The refractive index at a series of heights, based on data from the HESS site, is shown \ref{fig:lpd}. Exponentials are fitted between points to provide interpolation, and a comparison can be made to a linear interpolation that is also plotted. Despite the exponential interpolation, discontinuities in the second derivative of the refractive index prevent the LPD from being smooth in the simulation. This is unimportant, because in reality the variation due to atmospheric conditions and random noise will smear out any discontinuities in the LPD.

@@

\subsection{Classic HESS1 DC Pixel Identification}
We can find the DC pixel in an image with the variable $Q_{DC}$. In previous experiments, the DC pixel candidate was identified by applying a number of cuts to pixels in an image \cite{hess07}, and from the subset of pixels passing the specified cuts, selecting the pixel with the smallest $Q_{DC}$ as the \textquoteleft DC candidate'. Due to the low pass rate for cuts, we obtain a small low-contamination dataset, while the majority of telescope images are left without a DC pixel candidate.

As a basis for comparison, these original HESS cuts listed in Table \ref{tab:qdccuts} were replicated for the set of test data. For every HESS1 image, the total image amplitude $I_{tot}$ was used alongside the zenith angle $\theta$ to determine a dynamic cut, $Q_{DC} < 0.14 \times \log(\frac{I_{tot}}{161 \times \cos \theta})$. Because many images had no pixel that passed all cuts, the $Q_{DC}$ method was frequently unable to identify a DC pixel. In the original analysis, an additional cut $r_{core} \textgreater 40m$ was applied. However, the uncertainty in determining the core position through Hillas Analysis is typically of the order of $\pm 20m$. Consequently, this particular cut was omitted, though it would only serve to further reduce the DC pixel acceptance rate.

@@

In general, larger training sets prevent overtraining by making it more difficult to uniquely distinguish the individual events. The maximum depth of a tree is the number of branches a decision chain can pass through before it must end on a leaf. The minimum leaf sample is the minimum number that every leaf sample must exceed. Either of these quantities can also be used to restrict overtraining of a Decision Tree.

@@

The results are summarised in Table \ref{tab:qdcbdtcomparison1}. The BDT method represents a clear improvement in DC pixel identification over the previous $Q_{DC}$ method, and corresponds to an eightfold increase in identified pixels.

@@

\subsection{HESS-2 BDT}
The original HESS study predated the construction of the larger CT5 telescope, and thus focused exclusively in the four HESS-1 telescopes. Although the cuts in Table \ref{tab:qdccuts} were not optimised for the differing pixel size and angular viewing region of CT5, they were replicated to provide a basis for comparison. Based on CT5 images from the training sample, an increased 5.6\% of DC pixels were correctly identified and passed the required cut. A further 14.8 \% of all pixels passed the cuts, despite being misidentified. This represented a very heavy decrease in sample purity to just 27.5\%. We can assume that better optimised CT5 cuts could remove many of these misidentified events, though it is unlikely that any significant improvement in the number of correctly identified DC pixels would be possible. Thus, the number of DC pixels successfully identified using the $Q_{DC}$ method is still a valid benchmark for comparative BDT performance. A true accuracy rate closer to the HESS 1 rate of 80\% could be expected, at the expense of a depressed acceptance rate. Due to the distinctiveness of the CT5 telescope hardware, direct application of the HESS1 classifier to the CT5 telescope yields poorer BDT performance. To improve CT5 pixel identification, a separate HESS2 BDT was instead trained with the same variables as above. The relative feature importance is listed in Table \ref{tab:hess2classifier}.

@@

Despite the minor performance difference, the BDT method is clearly superior in both cases. On top the gains in BDT accuracy when considering only high-multiplicity events, we also find the relative performance gap over the $Q_{DC}$ method is vastly increased. As a result of the two-hundred-fold increase in acceptance rate for high-multiplicity-BDT events against high-multiplicity-$Q_{DC}$ events, the expected data sample size will increase dramatically. This should enable spectroscopic analysis of Cosmic Ray elements to be conducted after a much shorter experimental run time that for the $Q_{DC}$ method. Use of the BDT identification method will be assumed throughout the rest of this paper, and a high-multiplicity event will be intended to mean a high-multiplicity-BDT event.

\begin{table}[h!]
  \centering
  \caption{Comparison of high-multiplicity performance}
  \label{tab:highmultiplicitycomparison}
  \begin{tabular}{c|cc|cc}
    \toprule
    & \multicolumn{2}{c|}{HESS 1} & \multicolumn{2}{c}{HESS 2} \\
    & $Q_{DC}$ & BDT & $Q_{DC}$ & BDT\\
    \midrule
    High Multiplicity Accepted Pixels Correctly Identified (\%)& 0.01 & 2.20 & 0.01 & 1.61\\
    High Multiplicity Accepted Pixels Incorrectly Identified (\%)  & 0.00 & 0.43 & 0.01 & 0.22\\
    Sample Purity (\%)& 100.0 & 83.8 & 100.0 & 88.0\\
    \bottomrule
  \end{tabular}
\end{table}

@@

We require an additional cut to remove the majority of these proton events. If we can eliminate one or two accepted proton images per event, then the remaining proton pixels will all be rejected by the multiplicity cuts. Hence, loose cuts can sharply reduce the acceptance rate of protons. In this case, as shown in Figure \ref{fig:aspectratio}, there is a fair degree of separation between proton and Iron Aspect Ratio for pixels which pass all cuts. We aim to remove proton pixels but not iron pixels. Applying a cut requiring Aspect Ratio $> 0.40$ eliminated all accepted proton events for HESS1 and HESS2. With 18844 total HESS1 images, just 6195 images were triggered by DC light. Of all 18844 images, there were none passing all of the required cuts.

@@

Thus, if we have thirtyfold increase in event rate, but only one-fifteenth of the original acceptance rate, we would ultimately expect that we would have twice as many Iron Events if the full energy range of 5-135 TeV was included. However, the low-energy proton acceptance is much higher than before at 0.21\%. Thus, were we to include all events in this energy range, we would again retrieve a sample dominated by protons. The results are summarised in Table \ref{tab:fullironacceptance}.

@@

To remove the proton contamination from protons and non-saturated Iron events, we thus apply one additional Energy cut on the data sample. The majority of accepted protons have an Energy of less than 10 TeV, and the largest recorded Energy was 21.6 TeV. We can thus require that the reconstructed Energy is greater than 35TeV. This will, assuming a rough energy resolution of around 15\% in reconstruction, easily remove all of the proton background. We will also remove roughly half of the Iron events, but as shown in Figure \ref{fig:lpd}, this low-energy half of events will mostly contain non-saturated DC emission that would be harder to reconstruct.

@@

\section{Parameterisation of the LPD}
Having identified DC pixels in a way that is both reliable and consistently removes background events, we can now consider the resultant LPD. We must parameterise this function before we are able to reconstruct events. We can also make use of the additional information provided by the EAS light in each shower image.

\subsection{Full-Shower LPD}
Having restricted ourselves to high-multiplicity events, we reduce our data sample to mostly images in which air shower lies relatively close to the center of the telescope field of view. We can conclude that in many of our images, the air shower should be mostly contained within the telescope image. In the same way as for DC light, we can firstly look for the existence of a Characteristic LPD describing the EAS Intensity, which we define as $I_{tot}$=Image Amplitude. The results for high-multiplicity events at various energies is shown in Figure \ref{fig:fullshowerlpd}.

@@

For each energy, a clear exponential decay in Image Amplitude is observed. The data is fitted with a line of the form $y_{i} (x) =A_{i} \exp (K_{i} x)$. For the purposes of reconstruction, an exponential decay with a varying exponent and amplitude will not be particularly useful, because there are two degrees of freedom.

@@

Having retrieved a law for K, each dataset is then fitted a second time. The exponent is determined by the Energy, leaving only one degree of freedom for the fit. The fitted Amplitudes $A_{i}$ are then recorded, and are also plotted in Figure \ref{fig:powerlawlpd}. A clear relationship is again observed, this time an exponential law. A straight line is fitted to the logarithm of the data values, and we retrieve a second relationship describing the amplitude as a function of the Energy.

@@

Through combined use of the two equations, we can parameterise the entire Full-Shower LPD using only the Energy. These fitted LPDs are also plotted in Figure \ref{fig:fullshowerlpd}. The fractional deviation from the fit is found for each point by calculating $\Delta_{EAS} = \frac{fit_{EAS} - I_{tot}}{fit_{EAS}}$. We can then take the 68\% centile values to give us a $\sigma$-fraction width, and thus a fractional standard deviation $\frac{\sigma_{EAS}}{fit_{EAS}}$. The fractional standard deviation bands are shown in green, with a mean value of $\frac{\sigma_{EAS}}{fit_{EAS}}=0.15$ across the energies. We will make an approximation by assuming this value is a constant, though this is not strictly true. The Full Shower LPDs continue to be large enough to trigger telescopes for several hundred meters. It is thus likely that, for high-multiplicity events, we will obtain an additional four or five data points from the Full Shower LPD.

@@

Although the Full Shower LPDs do not depend upon the charge of the primary particle, they will still enable the core position and energy to be effectively constrained. We can use the Full Shower LPD to obtain a rough estimate for the core position, and thus a rough estimate of the distance from each telescope to the core. Similarly the core energy can be estimated. Although this reconstruction will be inferior to one accounting for both LPDs, we can make use of the preliminary core position estimate to aid our calculation of $True_{DC}$. The information from the full-shower LPD will also be incorporated into our final likelihood minimisation, when we reconstruct the events.

@@

\subsection{Determining $True_{DC}$}
Before we can extract the DC LPD from telescope images, we must first decide how to measure the quantity of DC light, $True_{DC}$. Determining the value of $True_{DC}$ is essential for determining the error in our LPD measurements. However, there are two natural ways of quantifying $True_{DC}$. In the basic case, we can simply consider the EAS-free intensity $Intensity_{max}$ in the DC pixel. An alternative is to take the entire EAS-free image amplitude $I_{tot}$ as our $True_{DC}$ value. The use of Image Amplitude accounts for the fact that the DC light is almost always split between two or more pixels. Although in most cases, one pixel has the majority of the light, in extreme cases the light can be evenly split between two. As a result of the smaller angular region covered by each CT5 pixel, this happens much more frequently for HESS2 than for HESS1. On the other hand, Image Amplitude measurement is susceptible to edge effects. A DC pixel lying near the image edge will be unaffected by this fact, but the total image amplitude may decline because dimmer pixels move over the image edge. We hope that, as with the full-shower LPD measurements, restricting ourselves to high-multiplicity events will enable us to remove most images that would otherwise suffer from edge effects.

To quantitatively compare the two methods, we can measure the error associated with both values of $True_{DC}$. A study of 2000 events was conducted, in which each cosmic ray was simulation without night sky background or EAS background. Each telescope image simulation was conducted twice, and difference between $I_{1}$ and $I_{2}$ was plotted as in Figure \ref{fig:simtelerror}. The fractional difference from the mean Intensity of an image was defined as $\Delta = 2 \times \frac{I_{2} - I_{1}}{{I_{2} + I_{1}}}$.
